---
title: "Task 4 analysis"
output: html_document
date: "2023-04-13"
---

```{r}
library(tidyverse)
library(here)
```

```{r}
read_csv(here("clean_data/candy_2015_to_2017.csv"))
```

### Question 1:
# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r}
# Specify the columns to include for counting candy ratings
candy_columns <- setdiff(colnames(candy_2015_to_2017), c("age", "year", "gender", "trick_or_treating", "country", "X"))

# Loop through the candy columns and count non-NA values
candy_counts <- numeric(length(candy_columns))
for (i in seq_along(candy_columns)) {
  candy_counts[i] <- sum(!is.na(candy_2015_to_2017[[candy_columns[i]]]))
}

# Create a data frame with column names and corresponding candy counts
candy_counts_df <- data.frame(Candy_Column = candy_columns, Count = candy_counts)

total_ratings <- candy_counts_df %>% 
  summarise(sum(Count))

total_ratings
```
 
# There are 598877	ratings 

### Question 2:
# What was the average age of people who are going out trick or treating?

```{r}
average_age_trick_or_treating <- candy_2015_to_2017 %>%
  filter(trick_or_treating == "Yes") %>%
  summarise(average_age = mean(age, na.rm = TRUE))

average_age_trick_or_treating 
```
# The average age of people going trick or treating is 35


## Question 3:
# What was the average age of people who are not going trick or treating?
```{r}
average_age_not_trick_or_treating <- candy_2015_to_2017 %>%
  filter(trick_or_treating == "No") %>%
  summarise(average_age = mean(age, na.rm = TRUE))

average_age_not_trick_or_treating 
```
# The average age of people NOT going trick or treating is 39

## Question 4:
#For each of joy, despair and meh, which candy bar received the most of these ratings
```{r}
candy_columns <- setdiff(colnames(candy_2015_to_2017), c("age", "year", "gender", "trick_or_treating", "country", "X"))

# Loop through the candy columns and count "joy", "despair", and "meh" ratings
joy_counts <- numeric(length(candy_columns))
despair_counts <- numeric(length(candy_columns))
meh_counts <- numeric(length(candy_columns))

for (i in seq_along(candy_columns)) {
  joy_counts[i] <- sum(candy_2015_to_2017[[candy_columns[i]]] == "JOY", na.rm = TRUE)
  despair_counts[i] <- sum(candy_2015_to_2017[[candy_columns[i]]] == "DESPAIR", na.rm = TRUE)
  meh_counts[i] <- sum(candy_2015_to_2017[[candy_columns[i]]] == "MEH", na.rm = TRUE)
}

# Create a data frame with column names and corresponding counts for "joy", "despair", and "meh" ratings
ratings_counts_df <- data.frame(Candy_Column = candy_columns, Joy_Count = joy_counts, Despair_Count = despair_counts, Meh_Count = meh_counts)

# Find the candy column with the most "joy" ratings
most_joy_candy <- ratings_counts_df %>%
  filter(Joy_Count == max(Joy_Count)) %>%
  select(Candy_Column)

# Find the candy column with the most "despair" ratings
most_despair_candy <- ratings_counts_df %>%
  filter(Despair_Count == max(Despair_Count)) %>%
  select(Candy_Column)

# Find the candy column with the most "meh" ratings
most_meh_candy <- ratings_counts_df %>%
  filter(Meh_Count == max(Meh_Count)) %>%
  select(Candy_Column)

cat("Candy bar with the most 'joy' ratings:", most_joy_candy$Candy_Column, "\n")
cat("Candy bar with the most 'despair' ratings:", most_despair_candy$Candy_Column, "\n")
cat("Candy bar with the most 'meh' ratings:", most_meh_candy$Candy_Column, "\n")

```

## Question 5: 
# How many people rated Starburst as despair?
```{r}
# Filter rows where candy column "starburst" has value "despair"
despair_starburst_count <- sum(candy_2015_to_2017$starburst == "DESPAIR", na.rm = TRUE)

# Display the count
cat("Number of people who rated starburst as 'despair':", despair_starburst_count, "\n")
```

## For the next three questions, count despair as -1, joy as +1, and meh as 0.

## Question 6: What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
# Filter out irrelevant columns
candy_data <- candy_2015_to_2017 %>%
  select(-year, -age, -trick_or_treating, -country, -X)

# Replace candy ratings with numeric values
rating_values <- c("JOY" = 1, "DESPAIR" = -1, "MEH" = 0)
for (col in colnames(candy_data)) {
  if (col != "gender") {
    candy_data[[col]] <- recode(candy_data[[col]], !!!rating_values)
  }
}

# Group by gender and calculate sum of ratings for all columns
candy_data <- candy_data %>%
  group_by(gender) %>%
  summarize(across(everything(), ~sum(., na.rm = TRUE)))

# Find the most popular candy bar for each gender
most_popular_candy_gender <- candy_data %>%
  pivot_longer(cols = -gender, names_to = "candy", values_to = "sum_ratings") %>%
  group_by(gender) %>%
  filter(sum_ratings == max(sum_ratings, na.rm = TRUE)) %>%
  select(gender, candy)

# Print the results
print(most_popular_candy_gender)
```
## Question 7: What was the most popular candy bar in each year?
```{r}
# Filter out irrelevant columns
candy_data <- candy_2015_to_2017 %>%
  select(-gender, -age, -trick_or_treating, -country, -X)

# Replace candy ratings with numeric values
rating_values <- c("JOY" = 1, "DESPAIR" = -1, "MEH" = 0)
for (col in colnames(candy_data)) {
  if (col != "year") {
    candy_data[[col]] <- recode(candy_data[[col]], !!!rating_values)
  }
}

# Group by year and calculate sum of ratings for all columns
candy_data <- candy_data %>%
  group_by(year) %>%
  summarize(across(everything(), ~sum(., na.rm = TRUE)))

# Find the most popular candy bar for each year
most_popular_candy_year <- candy_data %>%
  pivot_longer(cols = -year, names_to = "candy", values_to = "sum_ratings") %>%
  group_by(year) %>%
  filter(sum_ratings == max(sum_ratings, na.rm = TRUE)) %>%
  select(year, candy)

# Print the results
print(most_popular_candy_year)
```

## Question 8
```{r}
# Filter out irrelevant columns
candy_data <- candy_2015_to_2017 %>%
  select(-year, -age, -gender, -trick_or_treating, -X)

# Replace candy ratings with numeric values
rating_values <- c("JOY" = 1, "DESPAIR" = -1, "MEH" = 0)
for (col in colnames(candy_data)) {
  if (col != "country") {
    candy_data[[col]] <- recode(candy_data[[col]], !!!rating_values)
  }
}

# Group by country and calculate sum of ratings for all columns
candy_data <- candy_data %>%
  group_by(country) %>%
  summarize(across(everything(), ~sum(., na.rm = TRUE)))

# Find the most popular candy bar for each country
most_popular_candy <- candy_data %>%
  pivot_longer(cols = -country, names_to = "candy", values_to = "sum_ratings") %>%
  group_by(country) %>%
  filter(sum_ratings == max(sum_ratings, na.rm = TRUE)) %>%
  slice(1) %>%
  select(country, candy)

# Print the results
print(most_popular_candy)
```

