---
title: "Decathlon"
output: html_document
date: "2023-04-10"
---
```{r}
library(tidyverse)
library(here)
```

```{r}
decathlon <- read_csv(here("clean_data/decathlon.csv"))
```

## Question 1 
- Who has the longest jump? 
```{r}
long_jump_results <- decathlon %>% 
  select(athlete_name, long_jump, competition)
```

```{r}
arrange(long_jump_results, desc(long_jump)) %>% 
head(5)
```

Clay registered the longest jump at 7.96m at the Olympics 

##Question 2 
- What was the average 100m time in each competition?

```{r}
x100m_results <- decathlon %>% 
 group_by(competition) %>% 
  summarise(avg_100m_time = round(mean(x100m, na.rm = TRUE), 2))

x100m_results
```
The average time for Decastar is 11.18s and 10.92s for the Olympics 

## Question 3 
- Who has the highest points total across each competition 
```{r}
highest_points <- decathlon %>% 
  group_by(competition) %>% 
  filter(points == max(points)) %>% 
  select(competition, athlete_name, points)

highest_points
```

sebrle has the highest points total in each competition 

## Question 4 
- What was the shot-put scores for the top three competitors in each competition?

```{r}
top_3_shot_put_scores <- decathlon %>% 
  group_by(competition) %>% 
  arrange(desc(shot_put)) %>% 
  slice_head(n = 3) %>% 
  select(athlete_name, shot_put, competition)


top_3_shot_put_scores
```
For Decatsar the top 3 shot put scoreres were: 
1. yurkov
2. sebrel
3. karpov

For the olympics the top 3 shot put scoreres were: 
1. sebrle
2. karpov
3. macey

## Question 5 

```{r}
avg_points_400m_time <- decathlon %>%
  filter(x400m >= 50) %>%
  group_by(competition) %>%
  summarise(avg_points_above_50_runners = round(mean(points),2)) %>%
  inner_join(
    decathlon %>%
      filter(x400m < 50) %>%
      group_by(competition) %>%
      summarise(avg_points_sub_50_runners =round(mean(points),2)),
    by = "competition" 
  )

avg_points_400m_time
```

